Ecosystem Based Fisheries Management; Indicators; Maximum Sustainable Yield; Non-Stationarity; Productivity

### 2020-04-20 draft.tex
scrs version

\usepackage[authoryear]{natbib}
\usepackage{lineno}
\usepackage{pdfpages}
\usepackage{lscape}

@@

\title{Non-Stationarity in Productivity.}

@@

\abstract{
The Ecosystem Report Card of the Sub-Committee on Ecosystems includes indicators for assessed species  based on productivity, i.e. trends of biomass or spawning stock biomass relative and fishing mortality or harvest rate relative to Maximum Sustainable Yield reference points. The objective is to assess whether the main target stocks are in a healthy, cautious or critical state and how this has changed over time. Productivity, however, depends on physical and biological processes knowledge of which is important for ecosystem based fisheries management. We used the stock assessments for bigeye and yellowfin tunas from the Atlantic, Indian and Eastern Pacific Oceans, to illustrate the use of diagnostics based on production functions and surplus production verses biomass trajectories to explore changes in productivity.}

@@

Currently the main tool developed by the SCRS to implement EAFM is the Ecosystem Report Card (ERC) of the Sub-Committee on Ecosystems (SC-ECO). This includes indicators for assessed species and as the Commission moves towards EAFM there is a need to develop models for stock assessment and management that integrate environmental conditions \citep{travis2014integrating}.  Particularly since growth and mortality, and recruitment, may be driven by environmental pressure.

Currently the indicators used for assessed species are based on the ratios of biomass or spawning stock biomass (SSB) relative to $B_{MSY}$, and fishing mortality or harvest rate relative to $F_{MSY}$. These indicators are obtained either from age-structured stock assessments that integrate changing vulnerability schedules, mean recruitment relationships, and recruitment anomalies, to estimate stock trends and reference points, or biomass dynamic assessments that fit a production function based on population growth rate ($r$), carrying capacity ($K$). In the age based models there is an implicit production function and process error is modelled either as variability in recruitment and/or selection pattern. In the biomass dynamic models with an explicit production function process error can be modelled explicitly.

In age-structured models density dependence is mainly accounted for by the stock recruitment relationship. \cite{cury2014resolving}, however, showed that in most cases the stock-recruitment relationship used to estimate productivity and determine reference points, has poor estimation/predictive power and the environmental has a larger effect on productivity, a results confirmed by other studies \citep[e.g.][]{szuwalski2015examining, szuwalski2019global}(ADD Free, C.M., Thorson, J.T., Pinsky, M.L., Oken, K.L., Wiedenmann, J., Jensen, O.P., 2019. Impacts of historical warming on marine fisheries production. Sci 363, 979–983.) and observed 100 years ago by \cite{hjort1914fluctuations}. Whereas in ICCAT assessments growth, maturation and natural mortality are assumed not to have varied despite the large changes in the environment and stock biomass seen.

@@

@@

 Process error on biomass can account for model structural uncertainty as well as natural variability of stock biomass due to stochasticity in recruitment, natural mortality, growth, and maturation \citep{francis2011data} (add Meyer and Miller 1999; Winker et al. 2018).

@@

The observed dynamics may also be an indication of model misspecification. Diagnostic tools to check for model mis-specification and validation currently includes  deterministic Age-Structured Production Model (ASPM), likelihood profiling (Maunder and Piner xxx) and Runs Tests (Maunder & Piner 2015; Carvalho et al. 2017), but need to be developed further, specifically with regards to implication for assessment model prediction skill under non-stationary process error (Kell et al. 2016; Chang et al. 2019).

@@

@@

\section{Consequences for EBFM indicators}

In contrast Ecosystem models without age-structured effects typically predict the opposite pattern: higher SP at each stock size during recoveries than declines. Lower production during recoveries is often predicted by ecosystem models with size- and age-structured dynamics, especially when there are cultivation–depensation effects at play (Walters and Kitchell 2001, Liermann and Hilborn 1997).

Higher production rates during recoveries are sometimes predicted by ecosystem models with size- and age-structure effects, but only when such models include ecosystem-scale changes in productivity (regime shifts), such that trophically driven increases in production can drive population increases/vice versa where we don't observe these increases as a function of these changes in the ecosystem, and hence we see lower production at lower biomass levels.

Further, it has not been made clear what relationships we should expect between overall biomass and SP during population declines and recoveries and in particular whether age-structured and ecosystem models predict “non-stationarity” in the biomass–production relationship such that production rates exhibited at given biomass levels during recoveries might differ greatly from rates exhibited during declines. A particular concern is whether we should expect slower recoveries (lower SP at each stock size) than we would predict from production rates observed during declines, i.e., should populations recover more slowly than expected from simple dome-shaped production model theory (Hutchings and Reynolds 2004). For example, simulation studies have identified poor forecast performance when autocorrelation in recruitment was ignored (Johnson et al., 2016), which was associated with an increased risk of a “stochastic collapse” (Thorson et al., 2015).

In the Pacific Fishery Management Council (PFMC, Source , Citation), they use an age-structured model to predict recoveries so it would automatically incorporate at least recruitment effects caused by reduced mean fecundity in growing populations. Reduced mean fecundity in growing populations occurs because the population age structure is shifted toward younger fish, i.e., there is a proportional scarcity of older, more fecund individuals.

Here we examine observed relationships between SP rates and stock size for a variety of stocks. We estimate annual SP as change in stock size plus catch (SP = stock next year – stock this year + catch) and plot resulting empirical time series of SP and biomass estimates to look directly at patterns of variation in SP with biomass. We compare these patterns to those predicted by age-structured population models and by ecosystem models with and without age-structure effects.

We show that the usual assumptions in age-structured models (stable body growth, no change in natural mortality rate, recruitment following Beverton–Holt curve on average) generally imply lower SP during population recoveries than during declines because of reduction in mean fecundity. However, independent lognormal recruitment anomalies cause the opposite effect, with each positive anomaly resulting in a clockwise loop of successive values when SP is plotted against stock size; further, such anomalies can cause severe bias in estimates of unfished stock size, when a simple SP curve is fitted to the production vs. biomass data. We show that ecosystem models without age-structured effects typically predict the opposite pattern: higher SP at each stock size during recoveries than declines. Lower production during recoveries is often predicted by ecosystem models with size- and age-structured dynamics, especially when there are cultivation–depensation effects at play (Walters and Kitchell 2001).

Higher production rates during recoveries are sometimes predicted by ecosystem models with size- and age-structure effects, but only when such models include ecosystem-scale changes in productivity (regime shifts), such that trophically driven increases in production can drive population increases. We recommend that simple plots of SP rate vs. stock size be provided with stock assessment and ecosystem model results in general.

Winker et. al.(2020 jabbasel) noted that age-structured production models (ASPMs) are often preferred over biomass aggregated surplus production models (SPMs) as the former can track the propagation of cohorts and explicitly account for the effects of selective fishing, even in the absence of reliable size- or age data. ‘JABBA-Select’, an extension of the JABBA software (Just Another Bayesian Biomass Assessment; Winker et al., 2018), is able to overcome some of the shortcomings of conventional SPMs and allows a direct comparison to ASPMs.

\begin{itemize}[labelindent=\parindent,noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]
\item Finally, the presence of clockwise cycling due to recruitment anomalies, implies that future catches are driven by incoming year classes, rather than the production function.
\item There appears to be a suggestion that the stock may be more productive at low stock size that predicted by the implicit production function and reference points derived from them.

Additional work would need to look at estimates of growth and mortality, e.g. from tagging studies, to assess why these production functions are clockwise in direction.

As suggested earlier, there is possibility of model mis-specification, as some of the commissions omit the possibility of a SSR by setting the default a recruitment driven stock (h=0.99 in IATTC), which may or may not be the case.

Operating Models used in MSE should be conditioned on hypotheses related to ecological processes, particularly as this study showed that either the models used for stock assessment are mis-specified or do not account for important ecological processes.

\end{itemize}

@@

\clearpage
%\bibliography{/home/laurence-kell/Desktop/refs.bib}
%\bibliographystyle{apalike}
\input{draft.bbl}

>>>>

\usepackage[authoryear]{natbib}
%\usepackage{lineno}
\usepackage{pdfpages}
\usepackage{lscape}

@@

\title{Ecosystem Based Fisheries Management; Indicators; Maximum Sustainable Yield; Non-Stationarity; Productivity}

@@

\abstract{
The Ecosystem Report Card of the Sub-Committee on Ecosystems includes indicators for assessed species  based on productivity, i.e. trends of biomass or spawning stock biomass relative and fishing mortality or harvest rate relative to Maximum Sustainable Yield reference points. The objective is to assess whether the main target stocks are in a healthy, cautious or critical state and how this has changed over time. Productivity, however, depends on a variety of physical and biological processes. We therefore use the stock assessments for bigeye and yellowfin tunas from the Atlantic, Indian and Eastern Pacific Oceans, to illustrate the use of diagnostics based on production functions and surplus production trajectories to explore changes in productivity.}

@@

Currently the main tool developed by the SCRS to implement an EAFM is the Ecosystem Report Card (ERC) of the Sub-Committee on Ecosystems (SC-ECO). This includes indicators for assessed species and as the Commission moves towards EAFM there is a need to develop models for stock assessment and management that integrate environmental conditions \citep{travis2014integrating}.  Particularly since growth and mortality, and recruitment, may be driven by environmental pressure.

Currently the indicators used for assessed species are based on the ratios of biomass or spawning stock biomass (SSB) relative to $B_{MSY}$, and fishing mortality or harvest rate relative to $F_{MSY}$. These indicators are obtained either from age-structured stock assessments that integrate changing vulnerability schedules, mean recruitment relationships, and recruitment anomalies, to estimate stock trends and reference points, or biomass dynamic assessments that fit a production function based on population growth rate ($r$), carrying capacity ($K$). In the age-structured models there is an implicit production function and changes in productivity can occur due to process error modelled as variability in recruitment and/or selection pattern. In the biomass dynamic models with an explicit production function process error can be modelled explicitly.

In age-structured models density dependence is mainly accounted for by the stock recruitment relationship. \cite{cury2014resolving}, however, showed that in most cases the stock-recruitment relationship used to estimate productivity and determine reference points, has poor estimation/predictive power and the environmental has a larger effect on productivity, a results confirmed by other studies \citep[e.g.][]{szuwalski2015examining, szuwalski2019global,Free2019}, and observed 100 years ago by \cite{hjort1914fluctuations}. Whereas in ICCAT assessments growth, maturation and natural mortality are assumed not to have varied despite the large changes in the environment and stock biomass seen.

@@

%\citep{Maunder2015,Chang2019,Meyer1999,Thorson2015,Carvalho2017}

@@

 Process error on biomass can account for model structural uncertainty as well as natural variability of stock biomass due to stochasticity in recruitment, natural mortality, growth, and maturation \citep{francis2011data,Meyer1999,Thorson2015, Winker2018}.

@@

@@

The observed dynamics may be an indication of model misspecification. Diagnostic tools to check for model mis-specification and validation currently includes  deterministic Age-Structured Production Model (ASPM), likelihood profiling \citep{lee2014use} and Runs Tests \citep{Maunder2015,Carvalho2017}, but need to be developed further, specifically with regards to implication for assessment model prediction skill under non-stationary process error \citep{kell2016evaluation,Chang2019}.

@@

\section{Conclusions}

%In contrast Ecosystem models without age-structured effects typically predict the opposite pattern: higher SP at each stock size during recoveries than declines. Lower production during recoveries is often predicted by ecosystem models with size- and age-structured dynamics, especially when there are cultivation–depensation effects at play (Walters and Kitchell 2001, Liermann and Hilborn 1997).

%Higher production rates during recoveries are sometimes predicted by ecosystem models with size- and age-structure effects, but only when such models include ecosystem-scale changes in productivity (regime shifts), such that trophically driven increases in production can drive population increases/vice versa where we don't observe these increases as a function of these changes in the ecosystem, and hence we see lower production at lower biomass levels.

%Further, it has not been made clear what relationships we should expect between overall biomass and SP during population declines and recoveries and in particular whether age-structured and ecosystem models predict “non-stationarity” in the biomass–production relationship such that production rates exhibited at given biomass levels during recoveries might differ greatly from rates exhibited during declines. A particular concern is whether we should expect slower recoveries (lower SP at each stock size) than we would predict from production rates observed during declines, i.e., should populations recover more slowly than expected from simple dome-shaped production model theory (Hutchings and Reynolds 2004). For example, simulation studies have identified poor forecast performance when autocorrelation in recruitment was ignored (Johnson et al., 2016), which was associated with an increased risk of a “stochastic collapse” (Thorson et al., 2015).

%In the Pacific Fishery Management Council (PFMC, Source , Citation), they use an age-structured model to predict recoveries so it would automatically incorporate at least recruitment effects caused by reduced mean fecundity in growing populations. Reduced mean fecundity in growing populations occurs because the population age structure is shifted toward younger fish, i.e., there is a proportional scarcity of older, more fecund individuals.

%Here we examine observed relationships between SP rates and stock size for a variety of stocks. We estimate annual SP as change in stock size plus catch (SP = stock next year – stock this year + catch) and plot resulting empirical time series of SP and biomass estimates to look directly at patterns of variation in SP with biomass. We compare these patterns to those predicted by age-structured population models and by ecosystem models with and without age-structure effects.

%We show that the usual assumptions in age-structured models (stable body growth, no change in natural mortality rate, recruitment following Beverton–Holt curve on average) generally imply lower SP during population recoveries than during declines because of reduction in mean fecundity. However, independent lognormal recruitment anomalies cause the opposite effect, with each positive anomaly resulting in a clockwise loop of successive values when SP is plotted against stock size; further, such anomalies can cause severe bias in estimates of unfished stock size, when a simple SP curve is fitted to the production vs. biomass data. We show that ecosystem models without age-structured effects typically predict the opposite pattern: higher SP at each stock size during recoveries than declines. Lower production during recoveries is often predicted by ecosystem models with size- and age-structured dynamics, especially when there are cultivation–depensation effects at play (Walters and Kitchell 2001).

%Higher production rates during recoveries are sometimes predicted by ecosystem models with size- and age-structure effects, but only when such models include ecosystem-scale changes in productivity (regime shifts), such that trophically driven increases in production can drive population increases. We recommend that simple plots of SP rate vs. stock size be provided with stock assessment and ecosystem model results in general.

%Winker et. al.(2020 jabbasel) noted that age-structured production models (ASPMs) are often preferred over biomass aggregated surplus production models (SPMs) as the former can track the propagation of cohorts and explicitly account for the effects of selective fishing, even in the absence of reliable size- or age data. ‘JABBA-Select’, an extension of the JABBA software (Just Another Bayesian Biomass Assessment; Winker et al., 2018), is able to overcome some of the shortcomings of conventional SPMs and allows a direct comparison to ASPMs.

The presence of clockwise cycling due to recruitment anomalies, implies that future catches are driven by incoming year classes possibly due to environmental drivers rather than a production function. This has consequences for management based on target and limit reference points, since it follows that future biomass trends can not be predicted from current biomass based on setting total allowable catches.

There is also the possibility of model mis-specification, as some of the tRFMOs omit the possibility of a stock recruitment relationship by setting as the default that fact that recruitment driven stock (i.e. h=0.99 in IATTC), which may or may not be the case. It is important therefore as part of stock assessment to adopt quality control procedures to diagnose and 
 facilitat the interpretation of model mis-specification \citep{maunder2017dealing}.

Additional work is required to validate models and in particular to identify model mis-specifucation evaluate whether there have been changes in growth, mortality and  habitat that might have impacted in productivity. In this regards tagging studies are extremely valuable and historic tagging data should be analysed to identify hypotheses that can be tested in the future if the Commission is to move towards an EAFM in a cost effective way.

%There appears to be a suggestion that the stock may be more productive at low stock size that predicted by the implicit production function and reference points derived from them.

Operating Models for use in Management Strategy Evaluation should be conditioned on hypotheses related to ecological processes, particularly as this study showed that either the models used for stock assessment are mis-specified or do not account for important ecological processes.

@@

\clearpage
\bibliography{/home/laurence-kell/Desktop/refs.bib}
\bibliographystyle{apalike}
%\input{draft.bbl}
